Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-03-04 paper/paper.tex
pass on intro
pass on eval

\usepackage{subcaption}
\usepackage{libertine}
\usepackage{lipsum}% http://ctan.org/pkg/lipsum
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{graphicx}
\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption
\usepackage{booktabs}
%% \usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
%\usepackage[caption=false, font=footnotesize]{subfig}
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=true,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}
\usepackage[skins]{tcolorbox}

@@

Intuitively, rulesets need to be constantly revised to protect a
system against attackers. Unfortunately, creating these rules is
time-consuming and error-prone. The language for describing rules is
non-trivial (see Section~\ref{sec:example-suricata-rules}) and testing
correctness of a rule requires data and careful inspection.

@@

\subsection{Step 2: Rule Creation}
\label{sec:minimization}

@@

\newcommand{\textRQthree}{How important are the benign and
  malicious traffic to \tname's performance?}
\vspace{0.2cm}
\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ3.}} \textRQthree\
\end{itemize}

@@

\setlength{\tabcolsep}{5pt}
\begin{table}[h!]
  \small
  \caption{\label{table:results}Weights}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    Opt. &
    -f2 &
    -f3 &    
    -f4 &
    -f5 \\ 
    Black Nurse & 2 & 2 & 2 & 2 & 7$\uparrow$ \\    
    Ping Scan & 2 & 2 & 2 & 2 &  8$\uparrow$\\
    SYN Flood & 2 & 2 & 2 & 2 & 2 \\
    UDP Teardrop & 4 & 4 & 4 & 4 & 4 \\
    %%UDP Flood & - & - & - & - & - \\

@@

@@

>>>>

\usepackage{subcaption}
\usepackage{libertine}
\usepackage{lipsum}% http://ctan.org/pkg/lipsum
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{graphicx}
\usepackage[compatibility=false]{caption}% http://ctan.org/pkg/caption
\usepackage{booktabs}
%% \usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepgfplotslibrary{statistics}
\usepackage{listings}
%\usepackage[caption=false, font=footnotesize]{subfig}
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=true,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}
\usepackage[skins]{tcolorbox}

@@

Intuitively, rulesets need to be continuously revised because new
vulnerabilities are discovered and attackers can take advantage of
them. Unfortunately, creating these rules is time-consuming and
error-prone. The language for describing rules is non-trivial (see
Section~\ref{sec:example-suricata-rules}) and testing correctness of a
rule requires data and careful inspection.

@@

\subsection{Step 2: Rules Creation}
\label{sec:minimization}

@@

\newcommand{\textRQthree}{How sensitive \tname\ is to the amount of
  positive examples?}
\vspace{0.2cm}
\begin{itemize}[leftmargin=*,label={}]
\item{\textbf{RQ3.}} \textRQthree\
\end{itemize}

@@

This section evaluates the importance of each proposed heuristic
function. More precisely, we evaluated the use of \tname{} without
each of those functions with the goal of checking whether or not each
of them contributes to the overall performance of the technique. We
conducted that experiment by assigning the weight of the corresponding
function to zero---effectively, disabling the function---and comparing
the result with \tname\ using default weights, obtained with the
procedure described on Section~\Fix{xxx}.

Table~\ref{table:weights} shows results. Column ``Name'' shows the
name of the attack and column ``All'' shows the result of \tname\ with
all heuristic functions enabled (same as column ``Rank (step 3)'' from
Table~\Fix{xxx}). The remaining columns show results supressing one
heuristic function. An up arrow indicates that, considering that
particular data point, the performance of \tname{} on that
configuration is worse when compared to \tname's default
configuration. A down arrow, appearing highlighted in the table,
indicates the contrary.

\setlength{\tabcolsep}{5pt}
\begin{table}[h!]
  \small
  \caption{\label{table:weights}Contribution of heuristic functions.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    All &
    -f2 &
    -f3 &    
    -f4 &
    -f5 \\ 
    Black Nurse & 2 & 2 & 2 & 2 & 7$\uparrow$ \\    
    Ping Scan & 2 & 2 & 2 & 2 &  8$\uparrow$\\
    SYN Flood & 2 & 2 & 2 & 2 & 2 \\
    UDP Teardrop & 4 & 4 & 4 & 4 & 4 \\
    %%UDP Flood & - & - & - & - & - \\

@@

Note that the number of up arrows is relatively high and the magnitude
of the degradation can be very high in some cases. Basebuilder and
Tomcat are two cases where results get much worse when supressing one
of the functions. Also note that the number of down arrows is
relatively small. Recall that the optimization of weights looks for
global optimum and cannot avoid those cases. Finally, note that every
function contributed. Some functions contributed in more cases than
others, namely \Fix{xxx}, and one of the functions, namely \Fix{xxx},
only contributed to the performance of \tname{} on attacks involving
non-http protocols.

\begin{center}
\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
..........................
\end{tcolorbox}
\end{center}

@@

This section evaluates sensitivity of \tname{} to the number of
attacks provided on input. Recall that \tname\ uses one example attack
to generate the seed rule (at the Reverse Engineering step) and the
remaining attacks to evaluate soundness of the rule (at the Rules
Creations step). It is expected that the number and variety of
attacks---of the same kind---plays an important role in the
performance of the technique. To answer this research question, we
selected the cases with the highest number of examples of attacks (see
coolumn ``\#~Vars'' on Table~\ref{table:results}), varied the number of
examples, and measured the position of the golden rule at the
rank. For each number of variants $N$, we sampled five sets of size
$N$ to provide as input to the technique.
x

\Mar{coloquem os numeros aqui para que eu gere o boxplot amanha,
  depois da aula.

  *** posicao no ranking (numeros artificais. por favor, substituam) ***

  ataque 1:

    1 variacao:  10, 20, 4, 2, 5

    5 variacoes: 5, 4, 4, 2, 3

    10 variacoes: 1, 2, 1, 2, 1

    (ate numero de variacoes que aparece na Tabela 3)

  ataque 2:
     ...

%% \pgfplotsset{
%%   box plot width/.initial=0.5em
%% }
%% \begin{tikzpicture}
%%   \begin{axis}
%%     [
%%     ytick={1,2,3},
%%     yticklabels={Index 0, Index 1, Index 2},
%%     ]
%%     \addplot+[
%%     boxplot prepared={
%%       median=1,
%%       upper quartile=1.2,
%%       lower quartile=0.4,
%%       upper whisker=1.5,
%%       lower whisker=0.2
%%     },
%%     ] coordinates {};
%%     \addplot+[
%%     boxplot prepared={
%%       median=2,
%%       upper quartile=2.3,
%%       lower quartile=1.5,
%%       upper whisker=2.7,
%%       lower whisker=1
%%     },
%%     ] coordinates {};
%%     \addplot+[
%%     boxplot prepared={
%%       median=0.7,
%%       upper quartile=1.4,
%%       lower quartile=0.5,
%%       upper whisker=1.9,
%%       lower whisker=0.1
%%     },
%%     ] coordinates {};
%%   \end{axis}
%% \end{tikzpicture}
